---
title: "Data_analysis_birds"
output: 
  html_document:
    keep_md: true
---
# load required packages
```{r setup, include=FALSE}
pacman::p_load(tidyverse, broom, lubridate, gvlma, caret)
```

# Pull in clean data
```{r}
path <- "./Data_clean/"

files <- list.files(path=path, pattern="*.csv")

for(file in files){
  perpos <- which(strsplit(file, "")[[1]]==".")
  assign(
    gsub(" ","",substr(file, 1, perpos-1)),
    read_csv(paste(path,file,sep=""), guess_max = 10000))
}

```

## Join clean data
```{r}
full_data <- full_join(bird_diversity, bird_total) %>% 
  full_join(., daily_temp) 

View(full_data)
```

## Filter out NAs
```{r}
full_data <- full_data %>% 
  filter(!is.na(alpha & dailyMean))
```

# summarize data for statistics
```{r}
summary <- full_data %>% 
  group_by(siteID) %>% 
  summarize(n = n(),  # Sample size
            mean_dailyMean = mean(dailyMean), sd_dailyMean = sd(dailyMean), se_dailyMean = sd_dailyMean/sqrt(n), # Mean, sd, se for dailyMean
            mean_alpha = mean(alpha), sd_alpha = sd(alpha), se_alpha = sd_alpha/sqrt(n),
            mean_total = mean(total_birds), sd_total = sd(total_birds), se_total = sd_total/sqrt(n))

glimpse(summary)
```

## Round decimals
```{r}
summary <- as_tibble(summary) %>% 
  mutate_at(vars(mean_dailyMean:se_total), funs(round(.,2))) # Round each of our calculated columns to 2 decimals

print(summary)
```

# Linear regression: Alpha diversity
```{r Alpha_vs_Temp}
ggplot(data = full_data, aes(x = dailyMean, # name the variable that goes on the x-axis
                                    y = alpha,  # name the variable that goes on the y-axis
                                    col = siteID)) + # name the variable to differentiate colors
  geom_point(pch = 16) +  # Graph these variables as a scatterplot
  theme_minimal()
```

### Run correlation for diversity
```{r}
cor_div <- cor.test(full_data$dailyMean, 
                      full_data$alpha, 
                      method = "pearson") # can also use method = "spearman"

tidy(cor_div)
```

### Run linear regression test for diversity
```{r}
div_model <- lm(alpha ~ dailyMean, data = full_data)
```

#### Check model assumptions
```{r}
gvlma(div_model)
```

#### View model summary
```{r}
summary(div_model)
```

#### Re-graph data with regression line
```{r Alpha_vs_Temp_regressionLine}
ggplot(data = full_data, aes(x = dailyMean, # name the variable that goes on the x-axis
                                    y = alpha,  # name the variable that goes on the y-axis
                                    col = siteID)) + # name the variable to differentiate colors
  geom_point(pch = 16) +  # Graph these variables as a scatterplot
  geom_smooth(method = 'lm', aes(group=1), col = 'black', lty = 1) +  # lty=1 for solid line
  theme_minimal() +
  ggtitle("Alpha diversity vs Daily mean Temperature") +
  labs(x = "Daily mean Temp. (Celcius)", # Specify axis labels
       y = "Alpha diversity",
       fill = "siteID")
```

# Linear regression: Total birds
```{r Total_vs_Temp}
ggplot(data = full_data, aes(x = dailyMean, # name the variable that goes on the x-axis
                                    y = total_birds,  # name the variable that goes on the y-axis
                                    col = siteID)) + # name the variable to differentiate colors
  geom_point(pch = 16) +  # Graph these variables as a scatterplot
  theme_minimal()
```

### Run correlation for total
```{r}
cor_total <- cor.test(full_data$dailyMean, 
                      full_data$total_birds, 
                      method = "pearson") # can also use method = "spearman"

tidy(cor_total)
```

### run linear regression for total
```{r}
total_model <- lm(total_birds ~ dailyMean, data = full_data)
```
```{r}
gvlma(total_model)
```
```{r}
par(mfrow=c(2,2)) # init 4 charts in 1 panel
plot(total_model)
```

Failed linear regression assumption tests; check for and remove outlier values

## remove outliers then run regression of total
```{r}
subset <- full_data %>%
  filter(total_birds<300)
```

```{r}
cor_total_subset <- cor.test(subset$dailyMean, 
                      subset$total_birds, 
                      method = "pearson") # can also use method = "spearman"

tidy(cor_total_subset)
```
```{r}
total_model_subset <- lm(total_birds ~ dailyMean, data = subset)
```
```{r}
gvlma(total_model_subset)

```
```{r}
par(mfrow=c(2,2)) # init 4 charts in 1 panel
plot(total_model_subset)
```
```{r}
summary(total_model_subset)
```
```{r Total_vs_Temp_regressionLine}
ggplot(data = subset, aes(x = dailyMean, # name the variable that goes on the x-axis
                                    y = total_birds,  # name the variable that goes on the y-axis
                                    col = siteID)) + # name the variable to differentiate colors
  geom_point(pch = 16) +  # Graph these variables as a scatterplot
  geom_smooth(method = 'lm', aes(group=1), col = 'black', lty = 1) +  # lty=1 for solid line
  theme_minimal() +
  ggtitle("Total counts vs daily mean Temperature") +
  labs(x = "Daily mean Temp. (Celcius)", # Specify axis labels
       y = "Total bird count",
       fill = "siteID")
```

## Boxcox transformation
```{r}
distBCMod <- caret::BoxCoxTrans(full_data$total_birds, na.rm = T)
print(distBCMod)
```
```{r}
full_data <- cbind(full_data, total_new=predict(distBCMod, full_data$total_birds)) # append the transformed variable
head(full_data)
```
## Run anova test for total birds per site
```{r Total_boxplots}
ggplot(data = full_data, aes(x = siteID, # name the variable that goes on the x-axis
                                    y = total_birds,  # name the variable that goes on the y-axis
                                    fill = siteID)) + # name the variable to differentiate colors
  geom_jitter(width = 0.05, alpha = 0.6) +
  geom_boxplot(alpha=0.5) + # Graph these variables using boxplots
  theme_minimal() +
  ggtitle("Total count among sites") +
  labs(x = "Site ID", # Specify axis labels
       y = "Total bird count",
       fill = "Site")
```
## define ANOVA model
```{r}
total_ANOVA <- aov(total_birds ~ siteID, data = full_data)
summary(total_ANOVA)
```
```{r}
tukey_total <- tidy(TukeyHSD(total_ANOVA))
print(tukey_total)
```

## run ANOVA for alpha
```{r Alpha_boxplots}
ggplot(data = full_data, aes(x = siteID, # name the variable that goes on the x-axis
                                    y = alpha,  # name the variable that goes on the y-axis
                                    fill = siteID)) + # name the variable to differentiate colors
  geom_jitter(width = 0.05, alpha = 0.6) +
  geom_boxplot(alpha=0.5) + # Graph these variables using boxplots
  theme_minimal()  +
  ggtitle("Alpha Diversity among Sites") +
  labs(x = "Site ID", # Specify axis labels
       y = "Alpha diversity",
       fill = "Site")
```

```{r}
div_ANOVA <- aov(alpha ~ siteID, data = full_data)
summary(div_ANOVA)
```
```{r}
tukey_div <- tidy(TukeyHSD(div_ANOVA))

print(tukey_div)

tukey_div %>% 
  filter(adj.p.value < 0.05)
```

