---
title: "Data_wrangle_birds"
output: html_document
---
load required packages
```{r setup, include=FALSE}
#install.packages('pacman', repos = "http://cran.us.r-project.org")
pacman::p_load(tidyverse, lubridate)
```

pull in csv's for bird and temp data
```{r}
path <- "./Bird_data/"

files <- list.files(path=path, pattern="*.csv")

for(file in files){
  perpos <- which(strsplit(file, "")[[1]]==".")
  assign(
    gsub(" ","",substr(file, 1, perpos-1)),
    read_csv(paste(path,file,sep=""), guess_max = 10000))
}

path <- "./temperature_raw/"

files <- list.files(path=path, pattern="*.csv")

for(file in files){
  perpos <- which(strsplit(file, "")[[1]]==".")
  assign(
    gsub(" ","",substr(file, 1, perpos-1)),
    read_csv(paste(path,file,sep=""), guess_max = 500000))
}
```
Filter N/As out of temp data
```{r}
temp <- SAAT_30min %>% 
  filter(!is.na(tempSingleMean))
```

Remove columns not pertinent to our needs
```{r}
temp_sub <- temp %>% 
  mutate(date = as.Date(startDateTime),
         month = month(date),
         year = year(date),
         day = day(date),
         time = as_hms(startDateTime)) %>% 
  select(siteID, year, month, day, date, time, tempSingleMean:tempSingleMaximum, finalQF)

View(temp_sub)
```

Remove data that failed quality test
```{r}
temp_sub <- temp_sub %>% 
  filter(finalQF != 1)
```

Pull daily average temperature from the temp data
```{r}
daily_temp <- temp_sub  %>% 
  group_by(siteID, date) %>% 
  summarize(dailyMean = mean(tempSingleMean)) %>% 
  arrange(siteID, date)

View(daily_temp)
```

Write daily temp csv
```{r}
daily_temp %>% 
  write_csv('./Data_clean/daily_temp.csv')
```

change bird data name
```{r}
bird_raw <- brd_countdata
```
#select the columns we want to keep in our data
```{r}
bird_raw <- bird_raw %>%
  select(siteID, startDate, taxonID, vernacularName, clusterSize, taxonRank) %>%
  mutate(date = as.Date(startDate),
         time = as_hms(startDate))
         
```


#filter species out of taxonRank column
```{r}
bird_species <- bird_raw %>% 
  filter(taxonRank == "species")
```
#group by species
```{r}
bird_species_T <- bird_species %>%
  group_by(siteID, date, taxonID) %>%
tally()
```

#calculating alpha diversity

```{r}
bird_diversity <- bird_species_T %>%
  group_by(siteID, date)%>%
summarize(alpha = n_distinct(taxonID))
```
```{r}
bird_diversity %>%
  write_csv('./Data_clean/bird_diversity.csv')
```

eliminate NAs
```{r}
bird_count <- bird_raw %>% 
  filter(!is.na(clusterSize))
```


Find total bird count
```{r}
Bird_total<-bird_count %>%
  group_by(siteID, date) %>%
  summarize(total_birds = sum(clusterSize))
```

Write bird total csv
```{r}
Bird_total %>%
  write_csv('./Data_clean/bird_total.csv')
```